---
title: "R code for Mesopelagic role in the foodweb article"
output: html_notebook
---

Code to produce figures and analysis 


```{r Source functions, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

# set locale to avoid multibyte errors
Sys.setlocale("LC_CTYPE", "en_US.UTF-8")
# https://www.r-bloggers.com/web-scraping-and-invalid-multibyte-string/


setwd("~/gommesopelagic")

source("libraries.R")

source("map_polygons.R")
source("fishbase_data.R")
source("foodweb_diagram.R")
source("gom_diet_matrix.R")



```

Map of study area showing Atlantis polygons
```{r Make map, echo=TRUE, message=FALSE, warning=FALSE}
#scale factor is how much distance to add in the edges of the map, units are dependent on the coordinate system of the shapefile
#bar position can be top or bottom (t,b) or left or right (l,r)

model.map <- make_map(shape.file="GOM_LL.shp", file.name = "gom_model_map.png", scale.factor = 2, bar.position = "tl")

model.map

```

Retrieve habitat for species in each functional group to classify in guilds

```{r get_habitatdata, message=FALSE, warning=FALSE}

guild.names <- read_csv("fg_guild_names.csv")

atlantis.groups <- read_xlsx("Species_lists.xlsx", sheet = "group_names")

group.composition <- read_xlsx("Species_lists.xlsx", sheet = "group_species_composition") %>% 
  dplyr::rename(old_name=scientific_name)

species.list <- group.composition$old_name

fish.habitat <- lapply(species.list, get_fishbase) %>% 
  bind_rows

head(fish.habitat)

common.habitat <- fish.habitat %>% 
  left_join(group.composition, by="old_name") %>% 
  group_by(group,DemersPelag) %>%
  tally() %>% 
  top_n(1,n) %>% 
  mutate(habitat_classification = str_to_title(DemersPelag)) %>% 
  dplyr::select(-DemersPelag)
  
head(common.habitat)

common.habitat %>% 
  rename(long_name = group) %>% 
  right_join(guild.names, by="long_name") 

write_csv(common.habitat, "habitat_fish_categories.csv")


```

#foodweb diagram

```{r foodweb diagram, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
foodweb_map<- diet_plot(preynames="prey_names.csv", groupnames="Group_Names.csv", availabilitymatrix = "availabilities_matrix.csv", filename="gom_foodweb.png")

foodweb_map
```


Sourcing this file will extract mortality and growth scalars from the forcing file used to model the impacts of Deep Water Horizon in Ainsworth et al. 2018. for Small demersal fish, Deep water fish, and Other Demersal Fish.


```{r extract forcing, include=FALSE}

#needs original oil forcing files
source("modify_forcing.R")


```


Create pprey matrices using dummy availability values to test effects of contribution of mesopelagics on the food web

```{r dummy_availability, echo=FALSE}

workpath <- "~/gommesopelagic"

setwd(workpath)

guild.names <- read_csv("fg_guild_names.csv")

predator.guilds <- guild.names %>% 
  dplyr::rename(predator=fg_atlantis, predator_guild = guild) %>% 
  dplyr::select(predator, predator_guild)

prey.guilds <- guild.names %>% 
  mutate(`prey name` = tolower(long_name)) %>% 
  dplyr::rename(prey_guild = guild) %>% 
  dplyr::select(`prey name`, prey_guild)
  
prey.data.stat <-get_pprey(preynames = "prey_names.csv", functionalgroups="Group_Names.csv", availabilities = "availabilities_matrix.csv")

prey.guild <- prey.data.stat %>% 
  left_join(predator.guilds, by="predator") %>% 
  left_join(prey.guilds, by = "prey name") %>% 
  dplyr::select(-mean_av,-min_av,-max_av)

base.prey.guild.meso <- prey.guild %>% 
  filter((prey_guild=="Mesopelagic" & predator_guild=="Pelagic") & availability!=0)

base.prey.guild <- prey.guild %>%
  anti_join(base.prey.guild.meso, by=c("index","name","predator","predator name","predator type","prey type","prey name","predator_guild","prey_guild"))
  
#create logaritmic series

mult.seq <- lseq(0.00001,0.85,32)

savepath <- "~/gom_runs/runs_ppreyvalues"
headerfile  <- "GOM_PRM_2015_header.prm"
footerfile  <- "GOM_PRM_2015_footer.prm"
prey.frame.names  <- "prey_frame_names.csv"

mult.list <- 1:length(mult.seq)


lapply(mult.list,make_stanavail,base.prey.guild,base.prey.guild.meso,mult.seq,workpath,savepath,headerfile, footerfile, prey.frame.names)

sh.files <- 32

for(eachfile in 1:sh.files){

  file.rename(from=paste0("~/gom_runs/runs_ppreyvalues/goc_run_",eachfile,".sh"),to=paste0("~/gom_runs/runs_ppreyvalues/run_gom_oilmeso_dist_mpprey",eachfile,"/goc_run_",eachfile,".sh"))
  
  file.rename(from=paste0("~/gom_runs/runs_ppreyvalues/GOM_PRM_2015_",eachfile,".prm"),to=paste0("~/gom_runs/runs_ppreyvalues/run_gom_oilmeso_dist_mpprey",eachfile,"/GOM_PRM_2015_",eachfile,".prm"))
 
}    

lapply(mult.list,make_stanavail,base.prey.guild,base.prey.guild.meso,mult.seq,workpath,savepath,headerfile, footerfile, prey.frame.names)


for(eachfile in 1:sh.files){

  file.rename(from=paste0("~/gom_runs/runs_ppreyvalues/goc_run_",eachfile,".sh"),to=paste0("~/gom_runs/runs_ppreyvalues/run_gom_nooil_dist_mpprey",eachfile,"/goc_run_",eachfile,".sh"))
  
  file.rename(from=paste0("~/gom_runs/runs_ppreyvalues/GOM_PRM_2015_",eachfile,".prm"),to=paste0("~/gom_runs/runs_ppreyvalues/run_gom_nooil_dist_mpprey",eachfile,"/GOM_PRM_2015_",eachfile,".prm"))
} 


```



Run dummy availability values locally
```{r}
#used to remove extra forcing files
#list.files(pattern = "AtlantisSpatialForcing_1000Factor_363Thresh.nc",full.names = TRUE, recursive = TRUE) %>% grep("run_gom_oilmeso_dist_mpprey",., value=TRUE) %>% file.remove()

setwd("~/gom_runs/runs_ppreyvalues")

folder.paths <- c(paste0("run_gom_oilmeso_dist_mpprey",1:32),paste0("run_gom_nooil_dist_mpprey",1:32))
                  
folder.length <- 1:length(folder.paths)

system("sudo chmod -R a+rwx ~/gom_runs", wait = TRUE)

#run multiple Atlantis simulations on local machine cores

NumberOfCluster <- detectCores() # - 1

# Initiate cluster
cl <- snow::makeCluster(NumberOfCluster)
registerDoSNOW(cl)

# Run this for loop for one call of model from each cluster, assuming cluster is already initiated. 
atlantis.scenarios <- foreach(this.index=folder.length, .verbose = TRUE) %dopar% {
  
  # List of packages for session
  .packages = c("stringi","data.table","tidyverse","stringr","R.utils")
  
  # Install CRAN packages (if not already installed)
  .inst <- .packages %in% installed.packages()
  if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])
  
  # Load packages into session 
  lapply(.packages, require, character.only=TRUE)
  
  this.folder <- folder.paths[this.index]
  this.run <- this.folder %>% 
    str_split(.,"pprey") %>% 
    unlist %>% 
    .[2]
  
  setwd(paste0("~/gom_runs/runs_ppreyvalues/", this.folder))
  
  sh.file <- paste0("goc_run_",this.run,".sh")
  
  # run Atlantis scenario
  system(paste("sudo flip -uv *; sudo chmod +x ", sh.file,"; sudo sh ./", sh.file, sep=""), wait = TRUE)
  
  done <- as.data.frame("done")
}

stopCluster(cl)

system("az vm deallocate --name atlantisserver07 --no-wait --resource-group morzariacedogroup")


```

Analyze results from dummy availability values
```{r}

guild.names <- read_csv("fg_guild_names.csv") %>% 
  dplyr::rename(Code=group)

mult.seq <- lseq(0.00001,0.85,32)

biom.list <- list.files(path="~/gom_runs/runs_ppreyvalues",pattern="GOM_OUTBiomIndx.txt", recursive = TRUE, full.names = TRUE)

max.time <- 5474.5

baseline.data <- fread("~/gom_runs/gommodeloilmeso_revvertdist/output/GOM_OUTBiomIndx.txt",select= c(1:87)) %>% 
  pivot_longer(-Time, names_to = "species",values_to="biomass") %>% 
  filter(Time==max.time)

prey.meso.list <- grep("run_gom_oilmeso_dist_mpprey", biom.list, value=TRUE)

file.nums <- 1:length(prey.meso.list)

meso.data <- lapply(file.nums,get_biomass,prey.meso.list, max.time, this.scenario="meso_oil") %>% 
  bind_rows()

#Analyze no oil runs and pprey values

prey.nooil.list <- grep("run_gom_nooil_dist_mpprey", biom.list, value=TRUE)

file.nums <- 1:length(prey.nooil.list)

nooil.data <- lapply(file.nums,get_biomass,prey.nooil.list, max.time, this.scenario="no_oil") %>% 
  bind_rows() %>% 
  dplyr::rename(biomass_base = biomass)

ratio.data <- meso.data %>% 
  left_join(nooil.data, by=c("Time","species","preyvalue")) %>% 
  filter(Time==max.time) %>% 
  mutate(ratio = biomass/biomass_base) %>% 
  dplyr::rename(Code=species) %>% 
  left_join(guild.names, by="Code")

write_csv(ratio.data,"biomass_results.csv")

#smoothing from
#https://stats.idre.ucla.edu/r/faq/how-can-i-explore-different-smooths-in-ggplot2/
  
ratio.plot.meso <- ratio.data %>% 
  filter(guild=="Pelagic") %>% 
  filter(long_name %in% c("Large pelagic fish",
                            "Yellowfin tuna",
                            "Bluefin tuna",
                            "Other tuna",
                            "White marlin",
                            "Jacks",
                            "Large sharks")) %>% 
  ggplot(aes(x=preyvalue, y=ratio, color = long_name)) + 
  geom_point()+
#  facet_wrap(long_name ~., scales="free_y")+
  facet_wrap(long_name ~.)+
  theme_classic()+
  theme(legend.position = "none")+
  ylab("Biomass ratio") + 
  xlab("Prey availability value") +
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 1, colour="grey50")  #quadratic
 #  stat_smooth(method = 'lm', formula = y ~ poly(x,2), aes(colour = 'polynomial'), se= FALSE)+
 

ggsave("pprey_ratio_pelagic_plot.png", ratio.plot.meso, device="png",width=13, height = 11, units="cm")

ratio.plot.all <- ratio.data %>% 
  filter(!guild %in% c("Mesopelagic","Detritus","Bacteria",
                       "Plankton", "Primary producers","Infauna",
                       "Filter feeders","Epibenthos")) %>% 
   ggplot(aes(x=preyvalue, y=ratio, color = `guild`)) + 
  geom_point(colour = "black")+
  facet_wrap(long_name ~., scales="free_y")+
  theme_classic()+
  theme(legend.position = "none")+
  #theme(axis.text.y = element_text(angle = 90))+
  ylab("Biomass ratio") + 
  xlab("Prey availability value") +
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 1, colour="grey50")

ggsave("pprey_ratio_plot.png", ratio.plot.all, device="png",width=20, height = 10, dpi=300)


```

Analyze realized diets

```{r}

setwd("~/gom_runs/runs_ppreyvalues")

guild.names <- read_csv("fg_guild_names.csv") %>% 
  dplyr::rename(predator=fg_atlantis)

pel.groups <- c("Large pelagic fish",
                            "Yellowfin tuna",
                            "Bluefin tuna",
                            "Other tuna",
                            "White marlin",
                            "Jacks",
                            "Large sharks")

folder.paths <- list.dirs(full.names = TRUE, recursive = TRUE) %>% 
  grep(pattern="run_gom_*.*",., value = TRUE) %>% 
  grep(pattern="output_*.*",.,value = TRUE)
          
mult.seq <- lseq(0.00001,0.85,32)

col.names <-c("Time","prey","startN","mL","mQ","implicitMortality","GAG","RGR","SCM","SSR","DSR","RSN","VSN","LUT","BIO","LRF","SRF","BDR","RDR","SEA","SCI","LDY","MUL","POM","SHP","SNK","FLT","ODF","SDF","YTN","BTN","LTN","OTN","SWD","WMR","BMR","BIL","AMB","JCK","KMK","SMK","SAR","LPL","DWF","MEN","PIN","MPL","SPL","TIP","BEN","LGS","FIL","SMS","RAY","BSH","WSH","PSH","OSH","DBR","SBR","MAN","MYS","DOL","DDO","LOG","KMP","TUR","BCR","SCR","LOB","COR","CCA","OCT","SPG","CMB","INF","ECH","OYS","BIV","SES","EPI","GRS","ALG","MPB","LPP","SPP","DIN","PRO","JEL","SQU","LZP","SZP","PB","BB","DC","DL","DR","TotalPredMort")

base.prey <- get_mort(this.scenario = "~/gom_runs/gommodeloilmeso_revvertdist/output",mult.seq, col.names) %>% 
  dplyr::rename(base_end_mort=end_mortality, base_sum_end = sum_end_mort, base_prop_end = prop_end) %>% 
  dplyr::select(predator, prey, base_end_mort, base_sum_end, base_prop_end)

setwd("~/gom_runs/runs_ppreyvalues")

pprey.mort <- lapply(folder.paths, get_mort,mult.seq, col.names) %>% 
  bind_rows()


heatmap.prey <- pprey.mort %>% 
  left_join(base.prey, by = c("predator","prey")) %>% 
  left_join(guild.names, by=c("predator"))
  

write_csv(heatmap.prey,"~/gommesopelagic/prey_mortality.csv")

  
```


Plot prey mortality
```{r}

setwd("~/gommesopelagic")

guild.names <- read_csv("fg_guild_names.csv") %>% 
  dplyr::rename(prey=group, prey_name=long_name, prey_guild = guild)

heatmap.prey <- read_csv("prey_mortality.csv")

pel.groups <- c("Large pelagic fish",
                            "Yellowfin tuna",
                            "Bluefin tuna",
                            "Other tuna",
                            "White marlin",
                            "Jacks",
                            "Large sharks")

heatmap.data <- heatmap.prey %>% 
  filter(long_name %in% pel.groups) %>% 
  group_by(run, predator, availability) %>%
  mutate(cum_prop_end = cumsum(prop_end)) %>% 
  filter(prop_end != 0) %>% 
  filter(cum_prop_end > 0.30) %>% 
  ungroup %>% 
  left_join(guild.names, by="prey") %>% 
  mutate(prop_change = (((prop_end - base_prop_end)/base_prop_end) * 100)) %>% 
  filter(prey_guild %in% c("Demersal","Mesopelagic","Pelagic","Reef fish")) %>% 
  mutate(prop_change = if_else(is.infinite(prop_change),100,prop_change)) %>% 
  mutate(num_availability = dense_rank(availability))


mycols2 <- c("magenta3", #demersal
             "seagreen4", #mesopelagic
             "skyblue3", #pelagic
             "violetred4") # reef fish

myshapes <-c(15,17,19,8) 

prey.mortality <- ggplot(heatmap.data, aes(x=availability, y=prop_end, color=prey_guild, shape = prey_guild)) +
  geom_point() +
  scale_color_manual(values=mycols2, name="Prey guild") +
  scale_shape_manual(values=myshapes, name="Prey guild")+
  facet_wrap(long_name ~.)+
  theme_classic() +
  ylab("Prey mortality contribution") + 
  xlab("Prey availability")+
  theme(legend.position="bottom")

ggsave("pprey_mortality_plot.png", prey.mortality, device="png",width=14, height = 12, units="cm")


# ggplot(heatmap.data, aes(fill=prey_name, y=prop_change, x=long_name, group=num_availability)) + 
#    geom_bar(stat="identity", position = "dodge")+
#    theme_classic() +
#   coord_flip()+
#   facet_wrap(num_availability ~.)
   

# cl <- colors(distinct = TRUE)
# set.seed(345215) # to set random generator seed
# mycols2 <- sample(cl, length(unique(heatmap.data$prey_guild)))

# mycols2 <- c("tan4", #bacteria
#              "magenta3", #demersal
#              "orange", #epibenthos
#              "deepskyblue1", #filterfeeders
#              "dodgerblue4", #mesopelagic
#              "skyblue3", #pelagic
#              "darkseagreen", # plankton
#              "seagreen4", #primary producers
#              "cyan3", # reef fish
#              "violetred4", # sea turtles
#              "darkorchid2", # seabirds
#              "darkgoldenrod3") # squid



# ggplot(heatmap.data, aes(x=base_prop_end, xend=prop_end, y=availability, group=prey_guild)) + 
#   geom_dumbbell(point.colour.r="#a3c4dc", 
#                 size=0.75, 
#                 point.colour.l="#0e668b") + 
#   facet_wrap(long_name ~.)
# scale_x_continuous(label=prop_end)
# 
# ggplot(heatmap.data, aes(x=availability, y=prop_end, group=prey, color = prey_guild)) + 
#   geom_boxplot() +
#   scale_color_manual(values=mycols2) +
#   #  scale_color_brewer(palette = "Dark2")+
#   facet_wrap(long_name ~.)
# 
# ggplot(heatmap.data, aes(area = prop_end_per, fill = prey_guild, label = prey_guild)) +
#   geom_treemap() +
#   geom_treemap_text(fontface = "italic", colour = "white", place = "centre",
#                     grow = TRUE)+
#   facet_wrap(long_name ~.)
# 
# ggplot(heatmap.data, aes(prop_end, colour = long_name)) + stat_ecdf()+
#   facet_wrap(availability ~.)
# 
# # Plot again
# ggplot(heatmap.data, aes(x=availability, y=prop_end_per, fill=prey_guild)) + 
#   geom_area()+
#   scale_color_brewer(palette = "Dark2")+
#   facet_wrap(long_name ~.)
# 
# ggplot(heatmap.data, aes(fill=prey_guild, y=prop_end, x=availability)) + 
#   geom_bar(position="stack", stat="identity")+
#   theme_classic() +
#   facet_wrap(long_name ~.)
# 
# # filter(prop_end!=0) %>%   
# heatmap.data %>% 
#   ggplot(aes(predator,prey, fill = prop_end)) + 
#   geom_tile(aes(fill = prop_end)) + 
#   scale_fill_gradient(low = "lightblue", high = "firebrick4")+
#   theme_classic() +
#   facet_wrap(availability ~.)


```

Obtain catch data

```{r}
guild.names <- read_csv("fg_guild_names.csv") %>% 
  dplyr::rename(Code=group)

mult.seq <- lseq(0.00001,0.85,32)

catch.list <- list.files(path="~/gom_runs/runs_ppreyvalues",pattern="GOM_OUTCatch.txt", recursive = TRUE, full.names = TRUE)

max.time <- 5474.5

baseline.data <- fread("~/gom_runs/gommodeloilmeso_revvertdist/output/GOM_OUTCatch.txt",select= c(1:70)) %>% 
  pivot_longer(-Time, names_to = "species",values_to="catch") %>% 
  filter(Time==max.time)

prey.meso.list <- grep("run_gom_oilmeso_dist_mpprey", catch.list, value=TRUE)

file.nums <- 1:length(prey.meso.list)

meso.data <- lapply(file.nums,get_catch,prey.meso.list, max.time, this.scenario="meso_oil") %>% 
  bind_rows()

#Analyze no oil runs and pprey values

prey.nooil.list <- grep("run_gom_nooil_dist_mpprey", biom.list, value=TRUE)

file.nums <- 1:length(prey.nooil.list)

nooil.data <- lapply(file.nums,get_catch,prey.nooil.list, max.time, this.scenario="no_oil") %>% 
  bind_rows() %>% 
  dplyr::rename(catch_base = catch)

ratio.data <- meso.data %>% 
  left_join(nooil.data, by=c("Time","species","preyvalue")) %>% 
  filter(Time==max.time) %>% 
  mutate(ratio = catch/catch_base) %>% 
  dplyr::rename(Code=species) %>% 
  left_join(guild.names, by="Code")

write_csv(ratio.data,"catch_results.csv")

#smoothing from
#https://stats.idre.ucla.edu/r/faq/how-can-i-explore-different-smooths-in-ggplot2/
  

```

Plot catch data

```{r}

ratio.data <- read_csv("catch_results.csv") %>% 
  mutate(percent_change = ratio *100) 

ratio.plot.catch <- ratio.data %>% 
    filter(!guild %in% c("Marine mammals","Detritus","Bacteria",
                       "Plankton", "Primary producers","Infauna",
                       "Filter feeders","Seabirds","Sea turtles","Squid")) %>% 
  group_by(preyvalue, guild) %>% 
  summarise(tot_catch = sum(catch), tot_catch_base = sum(catch_base)) %>% 
  mutate(catch_ratio = tot_catch / tot_catch_base) %>% 
  #pivot_longer(cols=c("tot_catch","tot_catch_base"),names_to = "scenario") %>% 
   ggplot(aes(x=preyvalue, y=catch_ratio, color = guild)) + 
  geom_point()+
   facet_grid(guild ~ ., scales = "free_y")+
  theme_classic()+
  theme(legend.position = "none")+
  ylab("Catch ratio") + 
  xlab("Prey availability value") +
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 1, colour="grey50")  #quadratic
 #  stat_smooth(method = 'lm', formula = y ~ poly(x,2), aes(colour = 'polynomial'), se= FALSE)+
 

ggsave("pprey_ratio_catch_guild.png", ratio.plot.catch, device="png",width=13, height = 11, units="cm")


```


Plot ecosystem indices

Reef fish catch, Pelagic to demersal ratio, and Piscivore to planktivore ratio, and indicators robust to error, Red snapper biomass, Shrimp biomass, Forage fish biomass, and Elasmobranch biomass.

```{r}

#read biomass results

setwd("~/gommesopelagic")

biomass.data <- read_csv("biomass_results.csv")

catch.data <- read_csv("catch_results.csv") 

species.composition <- read_xlsx("Species_lists.xlsx", sheet = "group_species_composition") 

availability.data <- read_csv("prey_data_stat.csv")

atlantis.groups <- read_csv("Group_Names.csv")

guild.names <- read_csv("fg_guild_names.csv")

species.list <- species.composition %>% 
  pull(scientific_name)

fish.groups <- species.composition %>% 
  distinct(group) %>% 
  rename(long_name=group) %>% 
  left_join(guild.names, by = "long_name") %>% 
  pull(group)

plankton.groups <- c("LPP","SPP","DIN","PRO","JEL","SQU","LZP","SZP")

#run to get ecology data from Fishbase                    
#fish.ecology <- lapply(species.list, get_ecology) %>% 
#  bind_rows

#write_csv(fish.ecology, "fishbase_ecology.csv")

fish.ecology <- read_csv("fishbase_ecology.csv")
  
heatmap.prey <- read_csv("prey_mortality.csv")

fish.mort <- heatmap.prey %>% 
  filter(predator %in% fish.groups)

sum.fish.mort <- fish.mort %>% 
  group_by(predator) %>% 
  summarise(base_fin_mort=sum(base_end_mort))

fish.prey.mort <- atlantis.groups %>% 
  rename(`prey name`=`Long Name`, `prey` = Code) %>% 
  mutate(`prey name` = tolower(`prey name`)) %>% 
  right_join(fish.availability, by="prey") %>% 
  mutate(prey_category=if_else(prey %in% fish.groups,"fish prey",
                               if_else(prey %in% plankton.groups, "plankton prey","other prey")))

piscivore.code <- fish.prey.mort %>% 
    filter(prey_category=="fish prey") %>% 
    group_by(`predator`,prey_category) %>% 
    summarise(fish_prey_sum = sum(base_end_mort)) %>% 
    left_join(sum.fish.mort, by = c("predator")) %>% 
    ungroup %>% 
    mutate(fish_prey_ratio = fish_prey_sum / base_fin_mort) %>% 
    filter(fish_prey_ratio > 0.45) %>% 
    distinct(predator) %>% 
    pull(predator)

piscivore.groups <- atlantis.groups %>% 
  filter(Code %in% piscivore.code) %>% 
  transmute(long_name=tolower(`Long Name`)) %>% 
  pull(long_name)

planktivore.code <- fish.prey.mort %>% 
    filter(prey_category=="plankton prey") %>% 
    group_by(`predator`,prey_category) %>% 
    summarise(fish_prey_sum = sum(base_end_mort)) %>% 
    left_join(sum.fish.mort, by = c("predator")) %>% 
    ungroup %>% 
    mutate(fish_prey_ratio = fish_prey_sum / base_fin_mort) %>% 
    filter(fish_prey_ratio > 0.25) %>% 
    distinct(predator) %>% 
    pull(predator)

planktivore.groups <- atlantis.groups %>% 
  filter(Code %in% planktivore.code) %>% 
  transmute(long_name=tolower(`Long Name`)) %>% 
  pull(long_name)

pelagic.groups <- guild.names %>% 
  filter(habit=="Pelagic") %>% 
  transmute(long_name=tolower(long_name)) %>% 
  pull(long_name)

demersal.groups <- guild.names %>% 
  filter(habit=="Demersal") %>% 
  transmute(long_name=tolower(long_name)) %>% 
  pull(long_name)
  
reef.groups <- guild.names %>% 
  filter(guild=="Reef fish") %>% 
  transmute(long_name=tolower(long_name)) %>% 
  pull(long_name)

shrimp.groups <- c("White shrimp","Brown shrimp","Other shrimp","Pink shrimp") %>% 
  tolower()

forage.groups <- c("Small pelagic fish", "Medium pelagic fish") %>% 
  tolower()

elasmobranch.groups <- c("Filter feeding sharks","Large sharks","Small sharks","Benthic feeding sharks","Blacktip shark","Skates and rays") %>% 
  tolower()

avail.values <- biomass.data %>% 
  distinct(preyvalue) %>% 
  pull(preyvalue)

indicator.data <- lapply(avail.values, get_indicators, biomass.data,catch.data, reef.groups, pelagic.groups, demersal.groups, planktivore.groups, piscivore.groups, shrimp.groups, forage.groups, elasmobranch.groups) %>% 
  bind_rows

norm.indicator.data <- indicator.data %>% 
  group_by(availability) %>% 
  summarise(tot_oil_value = sum(oil_value), tot_base_value = sum(base_value)) %>% 
  left_join(indicator.data, by="availability") %>% 
  mutate(norm_oil_value = oil_value/tot_oil_value, norm_base_value = base_value/tot_base_value) 

#Radar plot
# norm.indicator.data %>% 
#   mutate(indicator=if_else(indicator=="Pelagic to demersal ratio","Pelagic:demersal ratio", 
#                            if_else(indicator=="Piscivore to planktonic ratio","Piscivore:planktonic ratio",indicator))) %>% 
#   dplyr::select(availability,indicator, norm_base_value, norm_oil_value) %>% 
#   filter(availability < 0.000011) %>% 
#   pivot_longer(norm_oil_value:norm_base_value,names_to = "scenario") %>% 
#   mutate(scenario = if_else(scenario=="norm_oil_value","Mesopelagic forcing","No oil")) %>% 
#   pivot_wider(id_cols=scenario,names_from = indicator, values_from = value) %>% 
#   ggRadar(., aes(group = scenario), 
#           rescale = FALSE, legend.position = "none",, colour = "darkslateblue",
#           size = 1, interactive = FALSE, use.label = TRUE) +
#   facet_wrap(~scenario)+
#   theme(axis.text.x = element_text(size = 8))

#lollipop plot for single availability
eco.inid.plot <- norm.indicator.data %>% 
  mutate(indicator=if_else(indicator=="Pelagic to demersal ratio","Pelagic:demersal ratio", 
                           if_else(indicator=="Piscivore to planktonic ratio","Piscivore:planktonic ratio",indicator))) %>% 
  dplyr::select(availability,indicator, norm_base_value, norm_oil_value) %>% 
  filter(availability < 0.000011) %>% 
  ggplot( aes(x=indicator, y=norm_base_value)) +
  geom_segment(aes(x=indicator, xend=indicator, y=norm_oil_value, yend=norm_base_value), color="darkgrey") +
  geom_point(size=5, color="darkcyan") +
  geom_point(aes(y=norm_oil_value), size=5, color="darkorange", alpha=0.7) +
  coord_flip() +
  theme_classic()+
  theme(legend.position = "none")+
  ylab("Value") + 
  xlab("Indicator") +
  ylim(0,1)

  ggsave("indicator_meso.png", eco.inid.plot, device="png",width=13, height = 11, units="cm")

  
  
eco.allav.plot <- norm.indicator.data %>% 
  mutate(indicator=if_else(indicator=="Pelagic to demersal ratio","Pelagic:demersal ratio", 
                           if_else(indicator=="Piscivore to planktonic ratio","Piscivore:planktonic ratio",indicator))) %>% 
  dplyr::select(availability,indicator, norm_base_value, norm_oil_value) %>% 
  ggplot( aes(x=indicator, y=norm_base_value)) +
  geom_segment(aes(x=indicator, xend=indicator, y=norm_oil_value, yend=norm_base_value), color="darkgrey") +
  geom_point(aes(color=availability),size=5, alpha=0.7) +
  geom_point(aes(y=norm_oil_value, color=availability), size=5, alpha=0.7) +
  coord_flip() +
  theme_classic()+
  ylab("Value") + 
  xlab("Indicator") +
  ylim(0,1)+
  theme(legend.position="bottom")+
  scale_color_continuous(name="Availability")


  ggsave("indicator_allmeso.png", eco.allav.plot, device="png",width=13, height = 11, units="cm")


```


