---
title: "R code for Mesopelagic role in the foodweb article"
output: html_notebook
---

Code to produce figures and analysis 


```{r Source functions, warning=FALSE, include=FALSE, paged.print=FALSE}

# set locale to avoid multibyte errors
Sys.setlocale("LC_CTYPE", "en_US.UTF-8")
# https://www.r-bloggers.com/web-scraping-and-invalid-multibyte-string/

#' LIBRARIES
#' -----------------
# List of packages for session
#if (!requireNamespace("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")

#BiocManager::install("affyio")

.packages = c("rgdal","data.table","tidyverse","here","maptools","broom","ggmap","rfishbase","devtools", "sf", "tmap","ggspatial","rgeos", "raster", "spData","spDataLarge","sp","rnaturalearth","rnaturalearthdata",
              "rnaturalearthhires", "readxl","scales","GGally","network","sna","RColorBrewer","grDevices",
              "colorRamps","colorspace","RNetCDF")

# Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])

# Load packages into session 
lapply(.packages, require, character.only=TRUE)

# if these packages are not present uncomment and run this code
# then uncomment before running chunk  
# remotes::install_github("ropensci/rfishbase")
#  install.packages('spDataLarge',
#repos='https://nowosad.github.io/drat/', type='source')
#  devtools::install_github("ropenscilabs/rnaturalearth")
#  devtools::install_github("ropenscilabs/rnaturalearthdata")
#  install.packages("rnaturalearthhires",
#                 repos = "http://packages.ropensci.org",
#                 type = "source")


lapply(.packages, require, character.only=TRUE)

source("map_polygons.R")
source("fishbase_data.R")
source("foodweb_diagram.R")

```

Map of study area showing Atlantis polygons
```{r Make map, echo=TRUE, message=FALSE, warning=FALSE}
#scale factor is how much distance to add in the edges of the map, units are dependent on the coordinate system of the shapefile
#bar position can be top or bottom (t,b) or left or right (l,r)

model.map <- make_map(shape.file="GOM_LL.shp", file.name = "gom_model_map.png", scale.factor = 2, bar.position = "tl")

model.map

```

Retrieve habitat for species in each functional group to classify in guilds

```{r get_habitatdata, message=FALSE, warning=FALSE}

atlantis.groups <- read_xlsx("Species_lists.xlsx", sheet = "group_names")

group.composition <- read_xlsx("Species_lists.xlsx", sheet = "group_species_composition") %>% 
  dplyr::rename(old_name=scientific_name)

species.list <- group.composition$old_name

fish.habitat <- lapply(species.list, get_fishbase) %>% 
  bind_rows

head(fish.habitat)

common.habitat <- fish.habitat %>% 
  left_join(group.composition, by="old_name") %>% 
  group_by(group,DemersPelag) %>%
  tally() %>% 
  top_n(1,n) %>% 
  mutate(habitat_classification = str_to_title(DemersPelag)) %>% 
  dplyr::select(-DemersPelag)
  
head(common.habitat)

write_csv(common.habitat, "habitat_fish_categories.csv")


```

```{r foodweb diagram, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
foodweb_map<- diet_plot(preynames="prey_names.csv", groupnames="Group_Names.csv", availabilitymatrix = "availabilities_matrix.csv", filename="gom_foodweb.png")

foodweb_map
```


Sourcing this file will extract mortality and growth scalars from the forcing file used to model the impacts of Deep Water Horizon in Ainsworth et al. 2018. for Small demersal fish, Deep water fish, and Other Demersal Fish.


```{r extract forcing, include=FALSE}

source("modify_forcing.R")

```

